---
title: "Open HDF5"
author: "Enrique Monta&ntilde;o"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning to load and work with HDF5 files in R

```{r load-libraries}
#load libraries
library(raster)
library(rhdf5)
library(rgdal)
```

## Open a file
```{r select-file}
# set the file path
f <- "../NEONdata/D17-California/TEAK/2013/spectrometer/reflectance/Subset3NIS1_20130614_100459_atmcor.h5"

# view h5 file structure
h5ls(f)

```

## Import Spatial Information

```{r import-spatial-info}
# import spatial information from the map info in the hdf file
mapInfo <- h5read(f, 
                  "map info",
                  read.attributes = TRUE)
# print the info
mapInfo
```

## Get the reflectance metadata

```{r get-refl-metadata}
# read the reflectance metadata attributes
reflInfo <- h5readAttributes(f,
                             "Reflectance")
# print the attributes
reflInfo

# extract the scale factor and no data value
scaleFactor <- reflInfo$`Scale Factor`
# check the structure
str(scaleFactor)

noDataValue <- reflInfo$`data ignore value`
str(noDataValue)

# promote the variable to numeric
noDataValue <- as.numeric(noDataValue)

```

## Import data dimensions

```{r import-dims}
# open file for viewing
fid <- H5Fopen(f) # open the file and assign file ID.  This is a reference only

# open the reflectance dataset
did <- H5Dopen(fid,
               "Reflectance") # open the dataset itself

 #view the attributes
did
# NOTE HDF is column-major, thus column/row.  R is row-major row/column
# The dims are listed as column x row x bands

# get the dataset dimensions
sid <- H5Dget_space(did)
sid

dims <- H5Sget_simple_extent_dims(sid)$size
dims #as column x rows x bands

# close all open connections
H5Sclose(sid) #dataspace
H5Dclose(did) #dataset
H5Fclose(fid) #file
```

## Read in reflectance data

```{r read-data}
# extract a slice of h5 file.  Band 56 in this case
# the rhdf library handles the indexing difference between the C=0 and R=1
b56 <- h5read(f,
              "Reflectance",
              index = list(1:dims[1], 1:dims[2], 56) )

class(b56)
```

## Convert the data to a matrix

```{r convert-matrix}
# convert to matrix
b56 <- b56[,,1] 
# using as.matrix(object) accomplishes the same thing

# let's plot some data
image(b56)

# stretch the image
image(log(b56), 
      main = "Log Transformed Data")

# look at why this worked
hist(b56)

# transformed - stretched
hist(log(b56))

```

## Time to clean up data
Apply the scale factor and no data values we obtained earlier.

```{r data-cleanup}
# Assign no data value
b56[b56 == noDataValue] <- NA

# apply scale factor
b56 <- b56 / scaleFactor
hist(b56, main = "Scaled data")


```

## Transpose data

We need to fix the orientation difference between column-major hdf and row-major R

```{r transpose-data}
# transpose the data
b56 <- t(b56)
image(log(b56), main = "Transposed Row/Column")
```

## Apply spatial extent

```{r apply-spatial}
# split the mapInfo object on the ","
mapInfo <- unlist(strsplit(mapInfo, ","))
# alternately, strsplit(mapInfo, ",")[[1]]
mapInfo

# define upper left coords
xMin <- as.numeric(mapInfo[4])
yMax <- as.numeric(mapInfo[5])

# define the resolution of the data
xRes <- as.numeric(mapInfo[6])
yRes <- as.numeric(mapInfo[7])

# define the upper right
xMax <- xmin + (dims[1] * xRes)
yMin <- yMax - (dims[2] * yRes)
```

## Create spatial extent

```{r create-extent}
# create a 
rastExtent <- extent(xMin, xMax,
                     yMin, yMax)

rastExtent

# create raster object
b56r <- raster(b56,
               crs = CRS("+init=epsg:32611"))
extent(b56r) <- rastExtent
b56r

plot(b56r,
     main = "Spatially referenced data")
```

## Import *NEON* Functions

```{r import-neon-functions}
# install devtools
# install.packages("devtools")
library(devtools)

# install_github("lwasser/neon-aop-package/neonAOP")
library(neonAOP)
# example using the NEON package
# use open_band function to view data
b55 <- open_band(f, 
                 bandNum = 55,
                 epsg = 32611)
b55

plot(b55, main = "NEON Function Loaded Data")


# we can import multiole bands at once
bands <- c(58, 34, 19) # red, green, blue

# create a raster stack
RGBStack <- create_stack(f,
                         bands = bands,
                         epsg = 32611)
RGBStack

# plot the stack as RGB
plotRGB(RGBStack,
        stretch = "lin")

# color IR
bands <- c(90, 34, 19)

# create a raster stack
CIRStack <- create_stack(f,
                         bands = bands,
                         epsg = 32611)

# plot the stack as RGB
plotRGB(CIRStack,
        stretch = "lin")



```

